---
title: "NYPD"
author: "Gia Long Tran"
date: "2023-09-17"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1:Import some libraries
```{r import_library}
library(tidyverse)
library(lubridate)

# define the dataset
url_in <- "https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD"

# Read data
Total_data <- read.csv(url_in)
```


## Step 2: Transform data(group by date, region, sex)

Read the data
```{r transform_data}
library(dplyr)

# Select relevant columns
new_data <- Total_data %>%
  select(OCCUR_DATE, OCCUR_TIME, BORO, PERP_SEX, VIC_SEX)

# Rename columns
colnames(new_data)[1] <- "Date"
colnames(new_data)[2] <- "Time"
colnames(new_data)[3] <- "Area"

# Convert Date column to date format
new_data <- new_data %>%
  mutate(Date = mdy(Date))

# Replace empty data with NA
new_data[new_data == ""] <- NA

```

## Step 3: Vizualize 2 graphs/charts

#### First visualization
Analyzing:We have the most incidents in Brooklyn area(over 10000 cases), and the least is contributed by Staten Island(around 800 cases)

```{r 1st_visualization, , echo=FALSE}
library(ggplot2)

# Create a bar plot
ggplot(new_data, aes(x = Area)) +
  geom_bar(fill = "red") +
  labs(title = "Incidents by Area", x = "Area", y = "Count")

```

#### 2nd visualization
Analyzing: From the chart, we can see male perpetrator is 10 times higher compare to female perpetrator(2500 versus 25000).

```{r 2nd_visualization, , echo=FALSE}
# Load necessary libraries
library(ggplot2)

# Create a bar chart of sex distribution
ggplot(data = new_data, aes(x = VIC_SEX)) +
  geom_bar(fill = "blue") +
  labs(title = "Sex Distribution by Perpetrator",
       x = "Sex",
       y = "Count") +
  theme_minimal()

```


## Step 4: modeling data
Using Time Series Model:  it is a set of data points over time and used in analysis to forcast the future

```{r}
library(ggplot2)
library(dplyr)

# Convert the "Date" column to a date format
new_data$Date <- as.Date(new_data$Date)

# Group data by date and count the number of incidents per day
incident_count <- new_data %>%
  group_by(Date) %>%
  summarize(Incidents = n())

# Create a time series plot
ggplot(incident_count, aes(x = Date, y = Incidents)) +
  geom_line() +
  labs(
    title = "Incidents Over Time",
    x = "Date",
    y = "Number of Incidents"
  ) +
  theme_minimal()
```

Analyzing: the time series model, we can see that number of incidents maintain quite stable over the years and different periods of time. 

## Bias indetifying
Based on the report data, we can assume that male can easily commit crime(shooting) comparing to female. Additionally, staying away from two areas Brooklynx and Bronx is highly recommended for better safety. Lastly, the highest number of incidents occur during 2020. Probably Covid-19 may have a significant impact into this data, which I believe mental health during quarantine time.
